A Machine-checked Proof of the Average-case Complexity of Quicksort in Coq

### 2008-08-29 paper/paper.tex
paper: separate abstract plus macro fiddling
abstract now comments out the bracket about translation between monoids;
comment macros now added, and a separate macros.sty file begun

darcs-hash:20080829140852-d98fc-d00f6165497425a064119bed20d3ce17fbbe15e7.gz

@@

\begin{abstract}
As a case-study in machine-checked reasoning about the complexity of algorithms in type theory, we describe a proof of the average-case complexity of Quicksort in Coq.
The proof attempts to follow a textbook development, at the heart of which lies a technical lemma about the behavior of the algorithm for which the original proof only gives an intuitive justification.
Our formalization includes a fully formal treatment of this lemma, introducing a number of technical innovations:
a shallow embedding of the algorithm as a monadically expressed functional program, instantiated with a variety of operation-counting monads that capture worst- and average case complexity of deterministic and non-deterministic programs, generalized to count in an arbitrary monoid (to support a key step in the proof where the expected comparison count is translated into the expected length of a recorded list of all comparisons);
a small theory of expectation for non-deterministically computed values, featuring both general map-fusion like results, and specific counting arguments.
\end{abstract}

>>>>

\usepackage{macros}

@@

\begin{abstract}
  \input{abstract.ltx}
\end{abstract}
